### 2014-11-04 thesis/method/evaluation.tex
Final version of thesis

The methods of evaluating the performance of classifiers in LOS prediction
that are discussed in the literature encompass a wide range of medical
domains and learning problems, which presents us with a challenge when we
decide how to evaluate our work. Here we have chosen to rigorously evaluate
our results (such as using stratified cross-validation when others may have
used a single training-test split), while still using performance measures
comparable to the literature and suitable for our particular learning problem.
This chapter discusses the procedure and performance measures used
in evaluating our work.

\section{Evaluation Procedure}
\subsection{Steps Involved}
The steps that we took to evaluate the performance of the various classifiers
and feature selection methods were:
\begin{enumerate}
\item The trauma LOS data was cleaned and preprocessed as described in Chapter
\ref{chap:preprocess}, giving us two data sets: one with and one without
discretisation. The general LOS data from Portugal was already cleaned when we
received it; however, we discretised the continuous \texttt{los} variable to
the same categories as for the trauma data (2 days or less, greater than 2
days), normalised the numeric attribute \texttt{prev\_admissions}, and derived
both a discretised and non-discretised data set. This resulted in 4 sets of
data.
\item For each data set, we evaluated the effect of the 4 automatic feature
selection methods on each of the 11 classifiers described in Chapter
\ref{chap:classification}, giving us 44 configurations for each data set, over
4 sets. Evaluation was performed using \textit{ten-fold stratified
cross-validation} (described below).
\item For the two trauma LOS data sets, we additionally evaluated the two
manual feature selection methods with all 8 classifiers. Recall from Section
\ref{sec:manual} that one of the feature sets is from the key previous work
on trauma LOS prediction that we use as a baseline, and the other feature set
is a set of features included through consultation with a domain expert.
\end{enumerate}

@@

\subsubsection{Code Written}
\label{sec:code}
Leading and trailing whitespace was removed and feature names transformed using
code written in Python (language version 3.4.1) to manipulate the CSV exported
from Microsoft Excel. All runs of WEKA were executed by
a program we wrote in Bash (a Unix shell) script. This program was able to run
WEKA with the following features:
\begin{itemize}
\item The ability to specify filenames corresponding to the input training data
and the specification of classification or feature selection algorithms to run
\item ROC curve data output can be optionally enabled and saved to disk after
each cross-validation pass
\item An experiment can be repeated an arbitrary number of times -- this is
specified by the user upon invoking the script (default is one time)
\item The ability to specify the inclusion of a random number seed -- this is
necessary for repetitions of runs to generate different cross-validation splits
\item Model information and performance measures are saved to disk after each
cross-validation
\end{itemize}

Because of the large numbers of results we obtained, we also wrote scripts in
Bash and Python to process and summarise the performance measures we required.
If we ran a set of classifiers 50 times in an experiment, the script would
retrieve all 50 outputs of a performance measure (such as accuracy, described
in the next section) for each classifier, and compute the mean and confidence
interval (also discussed in the next section) in a single output that we could
then save as CSV.

In order to implement our novel Ranked Distance Nearest Neighbour algorithm,
\ref{sec:rankeddist}), we created our own class within the WEKA Java package
hierarchy. This was implemented in Java and needed to implement the correct
interfaces and override the right inherited methods in order for the $k$-NN
classifier in WEKA to be able to use it as a distance function. This also
involved updating certain properties of the \texttt{weka.jar} executable in
order to use the new class, as well as building the entire WEKA source. The
development environment we used was Eclipse 4.2.1, and the
source was built using version 1.7.0 Update 17 of the Java Development Kit.

The source code listing of the most important programs we wrote are displayed
in Appendix \ref{app:code}.

@@

\subsection{Area Under the Curve}
The area under the receiver operating characteristic (ROC) curve, often
abbreviated AUC, is a measure of the ability of a binary classifier to
distinguish between members of the two classes. Let the \textit{true
positive rate} be $\dfrac{TP}{TP + FN}$ and the \textit{false positive rate}
be $\dfrac{FP}{FP + TN}$. The ROC curve is a plot of the true positive rate
on the vertical axis against the false positive rate on the horizontal axis.
The area under this curve will lie between 0 and 1, with 1 indicating a
100\% true positive rate and 0\% false positives (that is, everything was
classified correctly). The AUC of randomly picking between the two classes
with equal probability is 0.5, with a ROC curve that is a straight line
dividing the graph diagonally in half. Figure \ref{fig:sampleroc} shows
two ROC curves on the same graph:

@@

The AUC has not been commonly used in the literature to evaluate LOS
classification problems, but we have included it in our work for a number of
reasons. Firstly, it measures an aspect of classifiers that accuracy does not,
and which is more important and useful in the domain: the ability to
discriminate between instances of two classes. Recall from Chapter
\ref{chap:classification} that 0R will always predict the majority class --
this means that it will be evaluated positively on the basis of accuracy
as long as there are sufficiently more examples of one class than the other.
However, the AUC of the classifier will be \textit{less} than 0.5, because it
is less able to distinguish between the two classes than random guessing is,
due to the fact that it only ever predicts unseen examples to be in the
majority class. Secondly, previous work on trauma LOS prediction by Dinh et al.
also used the AUC as the key measure by which they evaluated their model
\cite{Dinh2013a}, and we would like our work to be comparable to theirs.

@@

\section{Summary}
In this section we described our method of evaluating the performance of each
classifier and feature selection method that we use. We used accuracy and the
area under the curve to measure two different aspects of our classifiers, and
we obtained these figures using stratified ten-fold cross-validation. Paired
$t$-tests were conducted to test for statistical significance of improvements
in accuracy and AUC.

>>>>

The methods of evaluating the performance of classifiers in LOS prediction
that are discussed in the literature encompass a wide range of medical
domains and learning problems, which presents us with a challenge when we
decide how to evaluate our work. Here we have chosen to rigorously evaluate
our results (such as using stratified cross-validation when others may have
used a single training-test split), while still using performance measures
comparable to the literature and suitable for our particular learning problem.
This chapter discusses the procedure and performance measures used
in our evaluation.

\section{Evaluation Procedure}
\subsection{Steps Involved}
The steps that we took to evaluate the performance of the various classifiers
and feature selection methods were:
\begin{enumerate}
\item The trauma LOS data was cleaned and preprocessed as described in Chapter
\ref{chap:preprocess}, giving us two data sets: one with and one without
discretisation. This was also done for the general hospital data set, resulting
in 4 sets of data.
\item For each data set, we evaluated the effect of the 5 automatic feature
selection methods on each of the 11 classifiers described in Chapter
\ref{chap:classification}, giving us 55 configurations for each data set, over
the 4 sets obtained in the previous step.
Evaluation was performed using \textit{ten-fold stratified
cross-validation} (described below).
\item For the two trauma LOS data sets, we additionally evaluated the two
manual feature selection methods with all 11 classifiers. Recall from Section
\ref{sec:manual} that one of the feature sets is from the key previous work
on trauma LOS prediction that we use as a baseline, and the other feature set
is a set of features derived through consultation with a domain expert.
\end{enumerate}

@@

\subsubsection{Code Written}
\label{sec:code}
Leading and trailing whitespace was removed and feature names transformed using
code written in Python (language version 3.4.1) to manipulate the CSV exported
from Microsoft Excel. All runs of WEKA were executed by
a program we wrote in Bash (a Unix shell) script, the source of which can be
found in Appendix \ref{sec:run-weka}. This program was able to run
WEKA with the following features:
\begin{itemize}
\item The ability to specify filenames corresponding to the input training data
and the specification of classification or feature selection algorithms to run
\item ROC curve data output can be optionally enabled and saved to disk after
each cross-validation pass
\item An experiment can be repeated an arbitrary number of times -- this is
specified by the user upon invoking the script (default is one time)
\item The ability to specify the inclusion of a random number seed -- this is
necessary for repetitions of runs to generate different cross-validation splits
\item Model information and performance measures are saved to disk after each
cross-validation
\end{itemize}

Because of the large numbers of results we obtained, we also wrote scripts in
Bash and Python to process and summarise the performance measures we required.
If we ran a set of classifiers 50 times in an experiment, the script would
retrieve all 50 outputs of a performance measure (such as accuracy, described
in the next section) for each classifier, and compute the mean and confidence
interval (also discussed in the next section) in a single output that we could
then save as CSV. These can be found in Appendix \ref{sec:summarise-script} and
\ref{sec:compute-ci}.

In order to implement our novel Ranked Distance Nearest Neighbour algorithm
(introduced in Section
\ref{sec:rankeddist}), we created our own class within the WEKA Java package
hierarchy. This was implemented in Java and needed to implement the correct
interfaces and override the right inherited methods in order for the $k$-NN
classifier in WEKA to be able to use it as a distance function. This also
involved updating certain properties of the \texttt{weka.jar} executable in
order to use the new class, as well as building the entire WEKA source. The
development environment we used was Eclipse 4.2.1, and the
source was built using version 1.7.0 Update 17 of the Java Development Kit.
The source for this can be found in Appendix \ref{sec:rankeddist-code}.

@@

\subsection{Area Under the Curve}
The area under the receiver operating characteristic (ROC) curve, often
abbreviated AUC, is a measure of the ability of a binary classifier to
distinguish between members of the two classes. Let the \textit{true
positive rate} be $\dfrac{TP}{TP + FN}$ and the \textit{false positive rate}
be $\dfrac{FP}{FP + TN}$. The ROC curve is a plot of the true positive rate
on the vertical axis against the false positive rate on the horizontal axis.
The area under this curve will lie between 0 and 1, with 1 indicating a
100\% true positive rate and 0\% false positives (that is, everything was
classified correctly). The AUC of randomly picking between the two classes
with equal probability is 0.5, with a ROC curve that is a straight line
dividing the graph diagonally in half. Figure \ref{fig:sampleroc} shows
three ROC curves on the same graph:

@@

The AUC has not been commonly used in the literature to evaluate LOS
classification problems, but we have included it in our work for a number of
reasons. Firstly, it measures an aspect of classifiers that accuracy does not,
and which is more important and useful in the domain: the ability to
discriminate between instances of two classes. Recall from Chapter
\ref{chap:classification} that Zero-R will always predict the majority class --
this means that it will be evaluated positively on the basis of accuracy
as long as there are sufficiently more examples of one class than the other.
However, the AUC of the classifier will be \textit{less} than 0.5, because it
is less able to distinguish between the two classes than random guessing is,
due to the fact that it only ever predicts unseen examples to be in the
majority class. Secondly, previous work on trauma LOS prediction by Dinh et al.
also used the AUC as the key measure by which they evaluated their model
\cite{Dinh2013a}, and we would like our work to be comparable to theirs.

@@

\section{Summary}
In this section we described our method of evaluating the performance of each
classifier and feature selection method that we used, including the code
written and software used. We used accuracy and the
area under the curve to measure two different aspects of our classifiers, and
we obtained these figures using stratified ten-fold cross-validation. Paired
$t$-tests were conducted to test for statistical significance of improvements
in accuracy and AUC.

### 2014-11-04 thesis/method/preprocess.tex
Final version of thesis

\section{Data Sets}
In our work we obtained two independent sets of LOS data, one specifically for
trauma and the other a general, hospital-wide data set.

@@

\subsubsection{Characteristics}
There were 2546 patient records in the data set we received, comprising of 79
features, one of which was the target variable \texttt{los48}.
\texttt{los48} was a binary variable -- that is, it could only take two
values, 0 or 1 -- with 1 indicating that the patient stayed two days or less,
and 0 indicating a stay of greater than 2 days.

@@

\section{Cleaning} % missing values, irrelevant features
In any data mining task, adequately \textit{cleaning} the data ensures the best
chance of success: this includes deciding how to deal with missing and
inaccurate values, as well as appropriately transforming the data
\citep{Witten2005}.
The exact steps involved in data cleaning will vary between data
sets, but the overall process is similar. Common issues to deal with include:
\begin{itemize}
\item Incorrect spelling or punctuation, perhaps as a result of mistakes in
data entry
\item Leading or trailing whitespace, and other garbage characters,
perhaps as a result of data entry or
the system that stores the data
\item Missing values for various features or for the class label
\item Values for features that do not make sense (such as a patient age of
-1)
\end{itemize}
The following paragraphs address how we dealt with the issues that were
applicable to our data.

@@

\paragraph{Extraneous whitespace or other characters.}
After correcting for spelling and punctuation errors,
the data was then exported into CSV (comma-separated value) format, so that
we could continue to process it programmatically rather than manually without
having to use Excel. Many
fields in the data set contained leading or trailing whitespace, which was
removed with a few lines of Python code (see Appendix
\ref{sec:python-preprocess}).
All attribute names were converted to lowercase. In the case
where attribute names consisted of several words, the spaces were replaced
with underscores (\_) to ensure that the full attribute name was read.
For example, \texttt{Any Cancer} was converted into \texttt{any\_cancer}.

@@

At the end of the cleaning process, the data was converted into the
Attribute-Relation File Format (ARFF)
\footnote{\url{http://www.cs.waikato.ac.nz/ml/weka/arff.html}} for use in
WEKA, a suite of machine learning algorithms implemented in Java
\cite{Hall2009}. Each feature was carefully checked to ensure that their
type was specified correctly -- that is, the nominal attributes were
distinguished from the numeric.

\subsection{General Hospital Data}
The general hospital data we received had already been preprocessed. The only
preprocessing here was to
undo the logarithm transform that had been applied to the \texttt{Previous
Admissions} feature.

\section{Normalisation}
Features represent different pieces of information about a training example.
Therefore, it is not surprising that not all features will have the same range
of values: a patient can have an \texttt{age} feature with a value of 20, but
their \texttt{heart rate} feature cannot possibly be 20. Learning algorithms
such as $k$-nearest neighbour (Chapter \ref{chap:classification})
are sensitive to the scales of feature values -- features with a larger range
of values will dominate a distance calculation. Therefore, we
\textit{normalise} all numeric (that is, continuous-valued)
feature values into the $[-1,1]$ range so that they are all expressed on the
same scale. Typically, values are normalised in the $[0,1]$ or the $[-1,1]$
ranges.
The formula to do this, also discussed in Chapter \ref{chap:classification},
is:
\begin{equation*}
a_i = 2\left[\dfrac{v_i - \mathrm{min }v_i}{\mathrm{max }v_i - \mathrm{min }v_i}\right] - 1
\end{equation*}
where we have used $s=2$ and $t=-1$ to transform values into the $[-1,1]$
range. This transformation was applied to each value of all numeric features
using WEKA, and was performed after all cleaning steps described above.

@@

Two basic approaches to the discretisation problem exist: unsupervised and
supervised. \textit{Unsupervised} discretisation involves splitting each
feature without considering any class information, whereas \textit{supervised}
discretisation accounts for the class when discretising.
Unsupervised discretisation uses simple methods such as splitting the data into
fixed-width intervals -- each interval spans the same width of values,
but can contain different numbers of feature values depending on the
distribution of underlying data -- which
can often partition the continuous feature into unhelpful categories due to its
ignorance of class labels. Supervised discretisation, on the
other hand, is commonly performed using the entropy minimisation heuristic
described by Fayyad and Irani \cite{Fayyad1993}. Since we have labelled
data, we are able to use both approaches. However, we will focus on evaluating
the effect of supervised discretisation, as it is generally preferred to
unsupervised discretisation when class information about the training examples
is known. And the method proposed by Fayyad and Irani is one of the best known
general supervised discretisation techniques \cite{Witten2005}.

\subsection{Entropy-based Discretisation}
In Chapter \ref{chap:classification} we outlined the induction of a decision
tree on a data set, using the information gain as the criterion for deciding
the attribute to split on at each step. A similar approach is used in Fayyad
and Irani's entropy-based method of discretisation \cite{Fayyad1993}:
to discretise an attribute,
each possible split point is considered and the information gain from that
split is computed. The split point with the highest information gain is
therefore chosen. Once the first split is determined, the algorithm continues
recursively in the newly created upper and lower ranges.

The \textit{minimum description length (MDL)} principle is used to determine
the stopping criterion. The MDL value is calculated by adding the information
values from two cases: split or do not split. If we split, then we can encode
the split point in $\mathrm{log}_2(N-1)$ bits (where $N$ is the number of
instances), then the classes of those above and below it. If we do not split,
then each instance's label must be encoded. Mathematically, given $N$ examples
with entropy $E$, sub-intervals $k_1$ and $k_2$ with the entropy of examples
in each sub-interval denoted $E_1$ and $E_2$:
\begin{equation*}
\mathrm{MDL} = \dfrac{\mathrm{log}_2(N-1)}{N} + \dfrac{\mathrm{log}_2(3^k-2) - kE + k_1E_1 + k_2E_2}{N}
\end{equation*}
The first term is the information needed for the splitting point, and the
second is a correction necessary for encoding which classes belong to which
sub-interval. The decision rule is then to split if the information gain for
that split exceeds the MDL value.

Since this requires exhaustive searching of the full range of feature values
to find potential split points, this method of supervised discretisation can
be time-consuming given large enough data sets. However, it has been proven
that any potential split that minimises the information value will never occur
between two examples of the same class. This means that we only need to
consider splits between regions of different classes, reducing the number of
possible points to consider.

@@

We then discussed normalisation, an important transformation of
continuous-valued data that ensures all features take on the same range of
values. This is important for ensuring that classifiers sensitive to the
relative range of values between features give valid outputs.

We concluded with outlining discretisation, in particular Fayyad and Irani's
entropy-based method. Such a technique has not been used in LOS prediction
and by including it in our work, we can evaluate its effectiveness in this
domain and let our results serve as a baseline and reference for future work.

>>>>

\section{Data Sets}
In our work we obtained two independent sets of LOS data through
collaboration with medical experts and researchers in Portugal: one
specifically for trauma and the other a general, hospital-wide data set.

@@

\subsubsection{Characteristics}
There were 2546 patient records in the data set we received, comprising of 79
features, one of which was the target variable \texttt{los48}.
\texttt{los48} was a binary variable -- that is, it could only take two
values, 0 or 1 -- with 1 indicating that the patient stayed two days or less,
and 0 indicating a stay of longer than 2 days.

@@

\section{Cleaning} % missing values, irrelevant features
In any data mining task, adequately \textit{cleaning} the data ensures the best
chance of success: this includes deciding how to deal with missing and
inaccurate values, as well as appropriately transforming the data
\citep{Witten2005}.
The exact steps involved in data cleaning will vary between data
sets, but the overall process is similar. Common issues to deal with include:
\begin{itemize}
\item Incorrect spelling or punctuation, perhaps as a result of mistakes in
data entry
\item Leading or trailing whitespace and other garbage characters,
perhaps as a result of data entry or the system that stores the data
\item Missing values for various features or for the class label
\item Values for features that do not make sense (such as a patient age of
-1)
\end{itemize}
The following paragraphs address how we dealt with the issues that were
applicable to our data.

@@

\paragraph{Extraneous whitespace or other characters.}
After correcting for spelling and punctuation errors,
the data was then exported into CSV (comma-separated value) format, so that
we could continue to process it programmatically, without the need for
Excel. Many fields in the data set contained leading or trailing whitespace,
which was removed with a few lines of Python code (see Appendix
\ref{sec:python-preprocess}).
All attribute names were converted to lowercase. In the case
where attribute names consisted of several words, the spaces were replaced
with underscores (\_) to ensure that the full attribute name was read.
For example, \texttt{Any Cancer} was converted into \texttt{any\_cancer}.

@@

At the end of the cleaning process, the data was converted into the
Attribute-Relation File Format (ARFF)
\footnote{\url{http://www.cs.waikato.ac.nz/ml/weka/arff.html}} for use in
WEKA, a suite of machine learning algorithms implemented in Java
\cite{Hall2009}. Each feature was carefully checked to ensure that its
type was specified correctly -- that is, the nominal attributes were
distinguished from the numeric.

\subsection{General Hospital Data}
The general hospital data we received had already been preprocessed. The only
preprocessing here was to
undo the logarithm transform that had been applied to the \texttt{Previous
Admissions} feature, because it had been previously applied during the course
of other work conducted on the data set. This was done using $e^x-1$, where
$x$ is the value of the feature.

\section{Normalisation}
Features represent different pieces of information about a training example.
Therefore, it is not surprising that not all features will have the same range
of values: a patient can have an \texttt{age} feature with a value of 20, but
their \texttt{heart rate} feature cannot possibly be 20. Learning algorithms
such as $k$-nearest neighbour (Chapter \ref{chap:classification})
are sensitive to the scales of feature values -- features with a larger range
of values will dominate a distance calculation. Therefore, we
\textit{normalise} all numeric (that is, continuous-valued)
feature values into the $[-1,1]$ range so that they are all expressed on the
same scale. Typically, values are normalised in the $[0,1]$ or the $[-1,1]$
ranges.
The formula to do this, also discussed in Chapter \ref{chap:classification},
is:
\begin{equation*}
a_i = 2\left[\dfrac{v_i - \mathrm{min }v_i}{\mathrm{max }v_i - \mathrm{min }v_i}\right] - 1
\end{equation*}
where we have used $s=2$ and $t=-1$ to transform values into the $[-1,1]$
range. This transformation was applied to each value of all numeric features
of both data sets
using WEKA, and was performed after all cleaning steps described above.

@@

Two basic approaches to the discretisation problem exist: unsupervised and
supervised. \textit{Unsupervised} discretisation involves splitting each
feature without considering any class information, whereas \textit{supervised}
discretisation accounts for the class when discretising.
Unsupervised discretisation uses simple methods such as splitting the data into
fixed-width intervals -- each interval spans the same width of values,
but can contain different numbers of feature values depending on the
distribution of underlying data -- which
can often partition the continuous feature into unhelpful categories due to its
ignorance of class labels. Supervised discretisation, on the
other hand, is commonly performed using the entropy minimisation heuristic
described by Fayyad and Irani \cite{Fayyad1993}. Since we have labelled
data, we are able to use both approaches. However, we will focus on evaluating
the effect of supervised discretisation, as it is generally preferred to
unsupervised discretisation when class information about the training examples
is known -- and the method proposed by Fayyad and Irani is one of the best
known general supervised discretisation techniques \cite{Witten2005}.

\subsection{Entropy-based Discretisation}
In Section \ref{sec:c45} we outlined the induction of a decision
tree on a data set, using the information gain as the criterion for deciding
the attribute to split on at each step. A similar approach is used in Fayyad
and Irani's entropy-based method of discretisation \cite{Fayyad1993}:
to discretise an attribute,
each possible split point is considered and the information gain from that
split is computed. The split point with the highest information gain is
therefore chosen. Once the first split is determined, the algorithm continues
recursively in the newly created upper and lower ranges.

The \textit{minimum description length (MDL)} principle is used to determine
the stopping criterion. The MDL value is calculated by adding the information
values from two cases: split or do not split. If we split, then we can encode
the split point in $\mathrm{log}_2(N-1)$ bits (where $N$ is the number of
instances) and subsequently the classes of those above and below it. If we do not split,
then each instance's label must be encoded. Mathematically, given $N$ examples
with entropy $E$, sub-intervals $k_1$ and $k_2$ with the entropy of examples
in each sub-interval denoted $E_1$ and $E_2$:
\begin{equation*}
\mathrm{MDL} = \dfrac{\mathrm{log}_2(N-1)}{N} + \dfrac{\mathrm{log}_2(3^k-2) - kE + k_1E_1 + k_2E_2}{N}
\end{equation*}
The first term is the information needed for the splitting point, and the
second is a correction necessary for encoding which classes belong to which
sub-interval. The decision rule is then to split if the information gain for
that split exceeds the MDL value.

Since this requires exhaustive searching of the full range of feature values
to find potential split points, this method of supervised discretisation can
be time-consuming given large enough data sets. However, it has been proven
that any potential split that minimises the information value will never occur
between two examples of the same class \cite{Witten2005}.
This means that we only need to
consider splits between regions of different classes, reducing the number of
possible points to consider.

@@

We then discussed normalisation, an important transformation of
continuous-valued data that ensures all features take on the same range of
values. This is important for ensuring that classifiers which are sensitive to
the relative range of values between features give valid outputs.

We concluded with outlining discretisation, in particular Fayyad and Irani's
supervised entropy-based method. Such a technique has not been used in LOS
prediction and by including it in our work, we can evaluate its effectiveness
in this domain and let our results serve as a baseline and reference for
future work.

### 2014-11-04 thesis/method/selection.tex
Final version of thesis

Our data set contains 78 features aside from the class, which is a fairly high
number. Many of these attributes
are derived from others in the set: for example, whether or not a patient is
over 65 years of age (the \texttt{age65} feature) is directly dependent on the
value of their \texttt{age} attribute.

@@

After eliminating redundant or inapplicable features as detailed above, we
considered four \textit{automatic} (correlation-based
feature selection, Pearson correlation, 1R and wrapper selection
using C4.5) and two \textit{manual} feature selection methods (domain expert
selection and the feature set used in existing work on trauma LOS prediction).
These methods are described in the following sections.

\section{Automatic Methods}
Automatic feature selection methods are useful for when we do not have a deep
understanding of the underlying domain and what each attribute really means to
the class. Even if we are well-versed in the domain, evaluating a few different
methods of automatic feature selection can give us new insight into the
learning problem at hand. Here we will describe a few of the methods we used.

@@

\subsection{1R}
Another way to evaluate the relevance of features is to use the idea behind the
1R classifier described in Section \ref{sec:zeror}. Recall that the 1R
classifier constructs simple 1-rules based on each feature and uses the feature
that produces the lowest error rate (that is, the highest accuracy) on the data
set to classify unseen examples. These accuracy rates can be calculated for
each feature, which are then used to rank the features from highest to lowest
accuracy. The worst accuracy for any feature is the 0R accuracy, which is the
proportion of the majority class in the data set. This is because if 1R finds
that the accuracy of the rule constructed using a feature is worse than 0R,
then it will not use that rule and predict the majority class instead.

@@

\subsection{Wrapper-based Feature Selection}
The feature selection methods described above (CFS, Pearson correlation and
1R) have been described as \textit{filter} approaches to attribute selection.
Kohavi and John proposed another approach, the \textit{wrapper} approach:
instead of performing feature selection independently of the learning
algorithm using some measure of ``relevance'',
the learning algorithm itself is used to evaluate the goodness of
subsets of features \cite{Kohavi1997}. The difference between the filter
and wrapper approaches is depicted in Figure \ref{fig:wrapper}.

@@

A key advantage of the wrapper approach is that the selected feature sets are
specifically tailored to a learning algorithm, which ensures the best
performance of that algorithm given the data set. However, wrapper approaches
can also be prohibitively slow if the data or feature sets are large and the
chosen learning algorithm expensive to train (such as support vector machines
and multi-layer perceptrons). We will evaluate the use of C4.5 decision trees
(see Section \ref{sec:c45}) in selecting features using the wrapper approach.
Using decision trees to select features is relatively fast as due to the
divide and conquer nature of the learning algorithm, and the tendency of the
trees to select fewer features for learning than other algorithms. Like CFS,
wrapper methods search for the best subset of features and do not require a
threshold to discard attributes.

\section{Manual Selection}
\label{sec:manual}
All of the above automatic feature selection methods use some notion of
``relevance'', such as symmetric uncertainty and linear correlation,
in order to determine the best subset of features to select.
Depending on the problem at hand, these definitions of relevance may or may
not work well with the data. In the LOS prediction literature, studies
have enlisted the help of medical domain experts to assist in feature
selection, due to their experience and understanding of the underlying
problem and the effect of the features on the outcome. It will therefore be
worthwhile to compare the performance of classifiers using features selected
automatically and manually.

@@

\subsection{Domain Expert Selection}
Consultation with a domain expert yielded another feature set that we could
use to compare with the baseline and with the other automatic methods.

>>>>

One of our data sets (the trauma-specific one) contains 78 features aside from
the class, which is a fairly large number. Many of these attributes
are derived from others in the set: for example, whether or not a patient is
over 65 years of age (the \texttt{age65} feature) is directly dependent on the
value of their \texttt{age} attribute. The other general hospital data set
contains 14 features, which is significantly less. For both data sets, we would
like to see if we can maintain or improve classifier performance by reducing
the number of features used.

@@

After eliminating redundant or inapplicable features as detailed above, we
considered five \textit{automatic} (correlation-based
feature selection, Pearson correlation, information gain, 1R and wrapper selection
using C4.5) and two \textit{manual} feature selection methods (domain expert
selection and the feature set used in existing work on trauma LOS prediction).
These methods are described in the following sections.

\section{Automatic Methods}
Automatic feature selection methods are useful when we do not have a deep
understanding of the underlying domain and the exact relationship of each
feature to the class. Even if we are well-versed in the domain, evaluating a
few different methods of automatic feature selection can give us new insight
into the learning problem at hand, and could also prove useful in extracting
new knowledge. Here we will describe a few of the methods we used.

@@

\subsection{1R}
Another way to evaluate the relevance of features is to use the idea behind the
1R classifier described in Section \ref{sec:zeror}. Recall that the 1R
classifier constructs simple 1-rules based on each feature and uses the feature
that produces the lowest error rate (that is, the highest accuracy) on the data
set to classify unseen examples. These accuracy rates can be calculated for
each feature, which are then used to rank the features from highest to lowest
accuracy. The worst accuracy for any feature is the Zero-R accuracy, which is
the proportion of the majority class in the data set. This is because if 1R
finds that the accuracy of the rule constructed using a feature is worse than
Zero-R, then it will not use that rule and predict the majority class instead.

@@

\subsection{Wrapper-based Feature Selection}
The feature selection methods described above (CFS, Pearson correlation and
1R) have been described as \textit{filter} approaches to attribute selection,
whereby the entire feature set is first ``filtered'' by some evaluation metric
and the reduced features used in a learning algorithm.
Kohavi and John proposed another approach, the \textit{wrapper} approach:
instead of performing feature selection independently of the learning
algorithm using some measure of ``relevance'',
the learning algorithm itself is used to evaluate the goodness of
subsets of features \cite{Kohavi1997}. The difference between the filter
and wrapper approaches is depicted in Figure \ref{fig:wrapper}.

@@

A key advantage of the wrapper approach is that the selected feature sets are
specifically tailored to a learning algorithm, which ensures the best
performance of that algorithm given the data set. However, wrapper approaches
can also be prohibitively slow if the data or feature sets are large and the
chosen learning algorithm expensive to train (such as support vector machines
and multi-layer perceptrons). We will evaluate the use of C4.5 decision trees
(see Section \ref{sec:c45}) in selecting features using the wrapper approach.
Using decision trees to select features is relatively fast as due to the
divide and conquer nature of the learning algorithm, and the tendency of the
trees to select fewer features for learning than other algorithms. Like CFS,
wrapper methods search for the best subset of features and do not require a
threshold to decide which features to select.

\section{Manual Selection}
\label{sec:manual}
All of the above automatic feature selection methods use some notion of
``relevance'', such as symmetric uncertainty and linear correlation,
in order to determine the best subset of features to select.
Depending on the problem at hand, these definitions of relevance may or may
not work well with the data. In the LOS prediction literature, studies
have enlisted the help of medical domain experts to assist in feature
selection, due to their experience and understanding of the underlying
problem and the effect of the features on the outcome. It will therefore be
worthwhile to compare the performance of classifiers using features selected
automatically and manually. We were able to obtain two sets of features
selected manually for the trauma data set.

@@

\subsection{Domain Expert Selection}
Consultation with a domain expert yielded another feature set that we
use to compare with the baseline and with the other automatic methods.
